---
title: "Correlation ParlVote"
output: html_notebook
---
The Notebook calculates the correlation between the Party Affiliation of the speaker and the Party bringing the motion in the ParlVote Corpus and compares this to the correlation between the predicted values of the lexica and the gold labels
```{r}
library(dplyr)
```


Make a Corpus which includes the party affiliation. 
```{r}
ParlVote_concat <- read.csv("~/Studium/3 Semester/PM/Corpora/ParlVote_concat.csv")
```

```{r}
#only use relevant columns and throw out entries with non ascii characters
parl_vote<- subset(ParlVote_concat, select = c("motion_party","party","speech","vote"))

parl_vote <-parl_vote[which(!grepl("[^\x01-\x7F]+",parl_vote$speech)),]

#Only choose the two biggest parties to make comparison easier
parl_vote<-subset(parl_vote, motion_party=='labour'|motion_party=="conservative")

parl_vote<-subset(parl_vote, party=='labour'|party=="conservative")


#Rename columns to standard column name
parl_vote<-parl_vote %>% 
 rename(
    text = speech,
    rating = vote)

#change the negative ratign to -1 to be consistent 
parl_vote$rating[parl_vote$rating == 0] <- -1


#sample 500 positive and negative examples
red_parl_vote <- data.frame(text=character(),
                 rating=integer(), 
                 stringsAsFactors=FALSE) 

hold<- subset(parl_vote, rating == -1)

red_parl_vote <- rbind(red_parl_vote, hold[sample(nrow(hold), 500), ])

hold<- subset(parl_vote, rating == 1)

red_parl_vote <- rbind(red_parl_vote, hold[sample(nrow(hold), 500), ])


#add ID column
red_parl_vote$id <- 1:nrow(red_parl_vote) 
red_parl_vote<-red_parl_vote[,c(5,1,2,3,4)]

```

```{r}
#save if needed
#saveRDS(red_parl_vote, file = "corr_part.rds")
```


Results of this corpus
```{r}
parl_res<- read.csv("~/Studium/3 Semester/PM/Datasets/parlvote_corr_norm1.csv")
```

Correlation between party and gold label
```{r}

corr_part<-parl_res[c("motion_party","party","rating")]

#Check if the speaker belongs to the party that brought up the motion
equ<-corr_part$motion_party==corr_part$party
corr_part$equality=equ


comp<-corr_part[c("rating","equality")]

party_sentiment_corr<-cor(comp$rating, comp$equality, method = c("pearson"), 
    use = "complete.obs")
```


```{r}
lex_corr<-parl_res[c("lsd","afinn","vader")]

#normalise the ratings to positive or negative
lex_corr[lex_corr>0] <- 1
lex_corr[lex_corr<=0] <- -1
 
#Choose one rating for each entry per majority vote 
lex_corr$sums=rowSums(lex_corr)

lex_corr[lex_corr>0] <- 1
lex_corr[lex_corr<=0] <- -1


lexicon_sentiment_corr<-cor(parl_res$rating, lex_corr$sums, method = c("pearson"), 
    use = "complete.obs")

```
